File-based localization of numerical perturbations in data analysis pipelines

### 2020-03-05 paper.tex
Update Figure-1
update example script, add new pg figures, add bin data samples

\begin{figure}
  \centering
    \includegraphics[width=1\columnwidth]{images/overview-fig}
    \caption{An overview of Spot tool in four steps.
    (\href{https://encsconcordiaca-my.sharepoint.com/:p:/g/personal/m_alari_encs_concordia_ca/EYx_aIZgYKlDuuXKPKi5ckIBo_WmJeE1Czr9rqG4gLHOOA?e=H8gack}{Source})}
    \label{fig:overview-tool}
  \end{figure}

@@

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}).

@@

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in purple and result files are in black.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. \todo{Tristan: check that, this is not a bipartite graph now}. 
}
  \label{fig:simple_script}
\end{figure}

@@

Files written by multiple processes are disambiguated as follows. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, it
 replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. \toolname finally
 updates the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1. This operation also makes the provenance graph
 acyclic, since we assumed that a process can only write a single time to a file.

As an example, in the sample pipeline in
Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a
temporary file. Also, file \emph{T1w\_brain\_bin.nii.gz} is first written
by \texttt{2\#bet2} and then overwritten by process \texttt{3\#fslmaths}
\todo{it's not the case, check}. The complete list of the processes that
write intermediary files is captured in Listing~\ref{listing:json_process}.
Using this list, we modify processes in Condition 1 and then re-execute
pipeline to save them in a backup folder.

\begin{listing}
  \begin{minted}[frame=single,
    framesep=3mm,
    linenos=false,
    xleftmargin=0pt,
    tabsize=4]{JSON}
{
  "multi_write_process": {
    "fslmaths\00T1w_brain.nii.gz\00T1w_brain_bin.nii.gz":
    {
      "files":[
        "/cond1/subj1/T1w_brain_bin.nii.gz"
      ],
      "pid": 78
    },
    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz":
    {
      "files":[
        "/cond1/subj1/T1w_brain_bin.nii.gz"
      ],
      "pid": 75
    }
  },
  "temp_process": {
    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz": 
    {
      "files": [
        "/cond1/subj1/temp.tmp"
      ],
      "pid": 75
    } 
  } 
}

\end{minted}
\caption{List of process and their files from the example pipeline in 
Listing~\ref{listing:sample-script} that should be captured.}
\label{listing:json_process}
\end{listing}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{images/iterative_modif}
  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red. \todo{update with bi-partite graph}}
  \label{fig:iterations}
\end{figure}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step one, process \texttt{1\#bet} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 2, \texttt{2\#bet2} 
is labeled as non-reproducible (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
Condition 1.
At step 3, \texttt{3\#fslstats} is labeled as 
reproducible as it writes files without differences.
At step 4, \texttt{4\#fslmaths} is labeled as non-reproducible 
as the last process of the pipeline.

>>>>

\begin{figure*}
  \centering
    \includegraphics[width=.8\textwidth]{images/spot-diagram}
    \caption{An overview of Spot tool in four steps.
    (\href{https://docs.google.com/drawings/d/1OB3sB8kkK17Q516-TcmXKh3TwXSLDToJB11IVfUH5AU/edit?usp=sharing}{Source})}
    \label{fig:overview-tool}
  \end{figure*}

@@

We use the \reprozip tool~\cite{rampin2016reprozip}
to capture: (1) the set of processes created by the
pipeline, and
(2) the set of files read and written by each process, including
temporary files. \reprozip collects this information through the
\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
\toolname then reconstructs a provenance graph by creating process and file
nodes from the \reprozip trace, and by adding directed edges corresponding
to file reads and file writes (Figure~\ref{fig:simple_script}-left).

@@

\begin{figure}
\centering
  \includegraphics[width=1\columnwidth]{images/provenance-graphs}
  \caption{Provenance graphs
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Nodes are devided into two sets including processes that are shaped 
  in the circles and files that are shaped in the boxes.
  Transient files are represented with grey color boxes. 
  Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies.
  Every node in the graph is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process or the file name.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  the rest of processes are the calls to FSL robustfov, bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Processes 75 and 149 are forked by processes 4 and 78, respectively: 
  they were captured by \reprozip while they did not appear in
  Listing~\ref{listing:sample-script}.  
}
  \label{fig:simple_script}
\end{figure}

@@

Files written by multiple processes are disambiguated as follows. For a
 file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
 $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
 write concurrently to F, which would violate our assumptions. Then, it
 replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
 calls $P_{i}$ and then saves F to a read-only directory. In this way,
 successive versions of F are preserved for comparison. \toolname finally
 updates the provenance graph accordingly, so that all files in the graph
 have an in-degree of 1 (Figure~\ref{fig:simple_script}-right). This operation also makes the provenance graph
 acyclic, since we assumed that a process can only write a single time to a file.

As an example, in the sample pipeline in
Listing~\ref{listing:sample-script}, process \texttt{3\#robustfov} writes a
temporary file. Also, file \emph{T1w\_brain.nii.gz} is first written
by \texttt{75\#bet2} and then overwritten by process \texttt{149\#bet2}. 
The complete list of the processes that
write transient files is captured in Listing~\ref{listing:json_process}.
Using this list, we modify processes in Condition 1 and then re-execute
pipeline to save them in a backup folder.

\begin{listing}
  \begin{minted}[frame=single,
    framesep=3mm,
    linenos=false,
    xleftmargin=0pt,
    tabsize=4]{JSON}
{
  "multi_write_process": {
    "bet2\00T1w\00T1w_brain":
    {
      "files":[
        "/cond1/subj1/T1w_brain.nii.gz"
      ],
      "pid": 75
    },
    "bet2\00T1w_roi\00T1w_brain":
    {
      "files":[
        "/cond1/subj1/T1w_brain.nii.gz"
      ],
      "pid": 149
    }
  },
  "temp_process": {
    "robustfov\00-i\00T1w.nii.gz\00-r\00T1w_roi.nii.gz": 
    {
      "files": [
        "/cond1/subj1/T1w_roi.nii.gz"
      ],
      "pid": 3
    } 
  } 
}

\end{minted}
\caption{List of processes and transient files from the example pipeline in 
Listing~\ref{listing:sample-script}.}
\label{listing:json_process}
\end{listing}

@@

\begin{figure*}
  \centering
  \includegraphics[width=.8\textwidth]{images/labelling-process}
  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
  \label{fig:iterations}
\end{figure*}

Figure~\ref{fig:iterations} illustrates our incremental labelling 
process for the example in Figure~\ref{fig:simple_script}. 
At step one, process \texttt{3\#robustfov} is executed and then labeled 
as reproducible (green) as it produced files without differences.
At step 2, \texttt{75\#bet2} 
is also labeled as reproducible as it produced files without differences. 
At step 3 \texttt{149\#bet2} is labeled as non-reproducible (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{149\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{149\#bet2} in 
Condition 1.
At step 4, \texttt{152\#fslmaths} is labeled as non-reproducible as it writes files with differences.
At step 5, \texttt{153\#fslstats} is labeled as 
reproducible as the last process of the pipeline.
